Individual heterogeneity in the functional topography of the medial prefrontal cortex

### 2019-03-14 Manuscript.Rmd
starting tsnr stuff

All network analyses were performed using the igraph package [@Csardi2006] in R 3.4.1 [@R Development Core Team]. To establish each subject's network, we selected all the vertices contained within the ROIs (n = 4,801 per subject) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, where surface vertices were the nodes, and edges the correlations among all of them. Next, we applied Fisher's r to z transformation, and performed a two-sided significance test to identify strong correlations. The resulting p-values were corrected for multiple comparisons (FDR < 0.05), and edges with non-significant correlations were turned to 0 (all other edges retained their respective z-values). Next, we took the exponential of these values so that all weights were positive while maintaining the ordinal ranks of the original correlation distribution. Values of 1 of this transformed distribution were returned to 0 in order to avoid the inclusion of these edges in the network analyses. We report the proportion of retained edges (i.e. network density) in the Results section. As part of the evaluation step, we produced these network weight matrices (and the community detection outlined below) at four levels: 1) for each run separately (1200 TRs each); 2) for the concatenated time series for the two runs on the same day (2400 TRs total); 3) for each subject's full dataset (4800 TRs total); and 4) on each step of a sliding window analysis (see Partition Evaluation for more details).

@@

# For each permutation, reassign the study labels, maintaining the original number of studies per condition
allStudies <- c(studyList_sv, studyList_dmn)
nPerms <- 100 # reported statistics are computed @ 5k, but it's computationally demanding to do that every time.
nullDist <- numeric()

for (i in seq(nPerms)) {

  # provide feedback only if requested (if TRUE, it will show up in the doc or pdf file)
  if (feedback) {
    if(i %in% c(nPerms*0.25, nPerms*0.5, nPerms*0.75, nPerms*0.95)) {print(paste("Completed", i/nPerms, "percent of the permutation..."))}
  }

  # Sample all the studies and divide into 2 groups
  randPerm <- sample(allStudies)
  tmpSV <- randPerm[seq(length(studyList_sv))]
  tmpDMN <- randPerm[-seq(length(studyList_sv))]

  # Count the number of times each region was reported in the literature
  tmpDF <- data.frame(Parcel = gsub("_","", Yeo_labels),
                      SV = sapply(Yeo_labels, function(roi) {length(grep(roi, tmpSV))}),
                      DMN = sapply(Yeo_labels, function(roi) {length(grep(roi, tmpDMN))}))

  # Perform the prop test for each pair and get the max chi-squared stat
  nullDist[i] <- max(
                  apply(tmpDF[,2:3], 1, function(data) {
                    prop.test(data, c(length(tmpSV), length(tmpDMN)))$statistic
                    }
                  ), na.rm = T)

# Now do the same for the actual data, to see how it compares to the null dist
tmpDF <- data.frame(Parcel = gsub("_","", Yeo_labels),
                    SV = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_sv))}),
                    DMN = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_dmn))}),
                    propSV = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_sv))}) / length(studyList_sv),
                    propDMN = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_dmn))}) / length(studyList_dmn))

# Perform the prop test for each pair and get the max chi-squared stat + pvalue based on the computed null distribution
tmpDF$Chi_squared <- apply(tmpDF[,2:3], 1, function(data) {
  round(prop.test(data, c(length(studyList_sv), length(studyList_dmn)))$statistic, digits = 2)
  })
tmpDF$Pval <- 1 - ecdf(nullDist)(tmpDF$Chi_squared) # probability of finding a chi-squared stat at least as large as the observed under a permuted null distribution

# ROIs that survived the threshold based on Chi-squared value (>95th quantile from null distribution -> 8.43)
significantROIs <- tmpDF %>% filter(Chi_squared >= quantile(nullDist, 0.95))

# results!
# 25 bilateral (prportion sv = L: 0.42, R: 0.39; proportion DMN = L: 0.18, R: 0.16, Chi = L: 12.91, R: 12.83, both p = 0.01)
# 7m bilateral (prportion sv = L: 0.17, R: 0.15; proportion DMN = L: 0.36, R: 0.40, Chi = L: 10.07, R: 18.89, L: p < 0.05, R: p < 0.0001)
# v23 right (proportion sv = 0.16; proportion DMN = 0.36, chi = 11.51, p = 0.01)
# null 95th-q chi <- 10.04 (varies; this was done ith 100 but resembles the 5k one)

@@

``` {r ROI computation, echo=FALSE, message = FALSE}
#TO DO: CHANGE THIS SO IT'S BASED ON STUDYLIST_SV/DMN, COUNTING JUST LIKE IN THE PERMUTATION ABOVE
## ROI definition
# First, load the raw count csv files
valCounts <- read.csv2('decCounts.csv', sep = ",")
dmnCounts <- read.csv2('dmnCounts.csv', sep = ",")

@@

Figure 1 shows the resulting ROIs from these analyses. The 95th percentile of the permuted Chi-squared distribution was `r quantile(nullDist, 0.95)`. Based on this threshold, area 7m was the only one to show unique affiliation to DMN (observed $\chi^2$ = 8.45, *p* < 0.05), while SV was uniquely represented by area 25 (observed $\chi^2$ = 11.95, *p* < 0.01). The purpose of these supra-threshold areas was to identify the most-likely affiliation of the two communities in our partitions by their coverage of these regions. However, area 25 sits in close proximity to the ventral striatum, which is reliably active in SV studies [@Bartra2013]. Considering that our projection from volumetric spheres to the cortex might have attributed striatal reports to this region, we decided not to use it as a marker of SV. We also chose not to use area v23 due to its unilateral significance. Instead, we only rely on area 7m to demarcate the DMN community.

@@

This comparison showed that both clustering methods were in high agreement across subjects (mean = `r round(mean(RI_all_within$Index), digits = 2)` sd = `r round(sd(RI_all_within$Index), digits = 2)`). Modularity showed a tendency to produce additional communities (median = `r median(nModularity)`, max = `r max(nModularity)`). However, these communities encompassed a small number of vertices (median = `r median(modularityCounts_extra, na.rm = T)`, IQR = `r IQR(modularityCounts_extra, na.rm = T)`) compared to the principal two (median = `r median(modularityCounts_two, na.rm = T)`, IQR = `r IQR(modularityCounts_two, na.rm = T)`), suggesting that the overall network is composed of two main communities. In addition, the estimation of the optimal number of cuts based on the rank of the first non-zero eigenvalue (see "Partition evaluation" for details) showed that `r prcntDissection` percent of the individual networks are best partitioned into two communities.

@@

``` {r Noise Effects: tSNR Comparison, echo=FALSE, message = FALSE, cache = TRUE}
## Putative steps to analyze the SNR of the full time series (PCC vs mPFC)
# Flexibility vs SNR
# Are vertices with lower SNR switching communities more often?
# This is pointless if thresholded maps are used, since flexibility is bound to be super low 
SNR_vs_Flex <- sapply(Summaries, function(data) cor(data$varCoeff, data$slideFlexibility))

# Compare SNR between PCC and mPFC per subject after thresholding (this filtering is relevant here because we aren't using ARI anymore)
tempComm_PCC <- lapply(tempComm_PCC, filter, FiedlerVec != 0)
tempComm_PFC <- lapply(tempComm_PFC, filter, FiedlerVec != 0) 
SNR_PCCvsPFC_perm <- sapply(seq_along(tempComm_PCC), function(x) {permute(tempComm_PCC[[x]]$varCoeff, tempComm_PFC[[x]]$varCoeff)$Pval})
SNR_PCCvsPFC_perm <- p.adjust(SNR_PCCvsPFC_perm, method = "BY")
SNR_PCCvsPFC_ES <- sapply(seq_along(tempComm_PCC), function(x) {cohenD(tempComm_PCC[[x]]$varCoeff, tempComm_PFC[[x]]$varCoeff)})
prcntSig <- mean(SNR_PCCvsPFC_perm < 0.05) * 100

@@

>>>>

All network analyses were performed using the igraph package [@G.2006] in R 3.4.1 [@RTeam2018]. To establish each subject's network, we selected all the vertices contained within the ROIs (n = 4,801 per subject) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, where surface vertices were the nodes, and edges the correlations among all of them. Next, we applied Fisher's r to z transformation, and performed a two-sided significance test to identify strong correlations. The resulting p-values were corrected for multiple comparisons (FDR < 0.05), and edges with non-significant correlations were turned to 0 (all other edges retained their respective z-values). Next, we took the exponential of these values so that all weights were positive while maintaining the ordinal ranks of the original correlation distribution. Values of 1 of this transformed distribution were returned to 0 in order to avoid the inclusion of these edges in the network analyses. We report the proportion of retained edges (i.e. network density) in the Results section. As part of the evaluation step, we produced these network weight matrices (and the community detection outlined below) at four levels: 1) for each run separately (1200 TRs each); 2) for the concatenated time series for the two runs on the same day (2400 TRs total); 3) for each subject's full dataset (4800 TRs total); and 4) on each step of a sliding window analysis (see Partition Evaluation for more details).

@@

# For each permutation, reassign the study labels, maintaining the original number of studies per condition
# allStudies <- c(studyList_sv, studyList_dmn)
# nPerms <- 100 # reported statistics are computed @ 5k, but it's computationally demanding to do that every time.
# nullDist <- numeric()
# 
# for (i in seq(nPerms)) {
# 
#   # provide feedback only if requested (if TRUE, it will show up in the doc or pdf file)
#   if (feedback) {
#     if(i %in% c(nPerms*0.25, nPerms*0.5, nPerms*0.75, nPerms*0.95)) {print(paste("Completed", i/nPerms, "percent of the permutation..."))}
#   }
#   
#   # Sample all the studies and divide into 2 groups
#   randPerm <- sample(allStudies)
#   tmpSV <- randPerm[seq(length(studyList_sv))]
#   tmpDMN <- randPerm[-seq(length(studyList_sv))]
# 
#   # Count the number of times each region was reported in the literature
#   tmpDF <- data.frame(Parcel = gsub("_","", Yeo_labels),
#                       SV = sapply(Yeo_labels, function(roi) {length(grep(roi, tmpSV))}),
#                       DMN = sapply(Yeo_labels, function(roi) {length(grep(roi, tmpDMN))}))
# 
#   # Perform the prop test for each pair and get the max chi-squared stat
#   nullDist[i] <- max(
#                   apply(tmpDF[,2:3], 1, function(data) {
#                     prop.test(data, c(length(tmpSV), length(tmpDMN)))$statistic
#                     }
#                   ), na.rm = T)
# 
# }
# 
# # Now do the same for the actual data, to see how it compares to the null dist
# tmpDF <- data.frame(Parcel = gsub("_","", Yeo_labels),
#                     SV = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_sv))}),
#                     DMN = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_dmn))}),
#                     propSV = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_sv))}) / length(studyList_sv),
#                     propDMN = sapply(Yeo_labels, function(roi) {length(grep(roi, studyList_dmn))}) / length(studyList_dmn))
# 
# # Perform the prop test for each pair and get the max chi-squared stat + pvalue based on the computed null distribution
# tmpDF$Chi_squared <- apply(tmpDF[,2:3], 1, function(data) {
#   round(prop.test(data, c(length(studyList_sv), length(studyList_dmn)))$statistic, digits = 2)
#   })
# tmpDF$Pval <- 1 - ecdf(nullDist)(tmpDF$Chi_squared) # probability of finding a chi-squared stat at least as large as the observed under a permuted null distribution
# 
# # ROIs that survived the threshold based on Chi-squared value (>95th quantile from null distribution -> 8.43)
# significantROIs <- tmpDF %>% filter(Chi_squared >= quantile(nullDist, 0.95))

# results!
# 25 bilateral (prportion sv = L: 0.42, R: 0.39; proportion DMN = L: 0.18, R: 0.16, Chi = L: 12.91, R: 12.83, both p = 0.01)
# 7m bilateral (prportion sv = L: 0.17, R: 0.15; proportion DMN = L: 0.36, R: 0.40, Chi = L: 10.07, R: 18.89, L: p < 0.05, R: p < 0.0001)
# v23 right (proportion sv = 0.16; proportion DMN = 0.36, chi = 11.51, p = 0.01)
# null 95th-q chi <- 9.11 (varies; this was done ith 100 but resembles the 5k one)

@@

``` {r ROI computation (overlapping: old), echo=FALSE, message = FALSE}
#TO DO: CHANGE THIS SO IT'S BASED ON STUDYLIST_SV/DMN, COUNTING JUST LIKE IN THE PERMUTATION ABOVE
## ROI definition
# First, load the raw count csv files
valCounts <- read.csv2('decCounts.csv', sep = ",")
dmnCounts <- read.csv2('dmnCounts.csv', sep = ",")

@@

Figure 1 shows the proportion of times every parcel was reported in each literature, as well as the resulting significant ROIs from the permutation. The 95th percentile of the permuted Chi-squared distribution was 9.11. Based on this threshold, area 7m was the only one to show unique bilateral affiliation to DMN (Left: observed $\chi^2$ = 10.07, *p* = 0.034; Right: observed $\chi^2$ = 18.89, *p* < 0.0001), while SV was uniquely represented bilaterally by area 25 (Left: observed $\chi^2$ = 12.91, *p* = 0.014; Right: observed $\chi^2$ = 12.83, *p* = 0.014). The purpose of these supra-threshold areas was to identify the most-likely affiliation of the two communities in our partitions by their coverage of these regions. However, area 25 sits in close proximity to the ventral striatum, which is reliably active in SV studies [@Bartra2013]. Considering that our projection from volumetric spheres to the cortex might have attributed striatal reports to this region, we decided not to use it as a marker of SV. We also chose not to use area v23 due to its unilateral significance (Right: observed $\chi^2$ = 11.51, *p* = 0.018). Instead, we only rely on area 7m to demarcate the DMN community.

@@

This comparison showed that both clustering methods were in high agreement across subjects (mean = `r round(mean(RI_all_within$Index), digits = 2)`, SD = `r round(sd(RI_all_within$Index), digits = 2)`). Modularity showed a tendency to produce additional communities (median = `r median(nModularity)`, max = `r max(nModularity)`). However, these communities encompassed a small number of vertices (median = `r median(modularityCounts_extra, na.rm = T)`, IQR = `r IQR(modularityCounts_extra, na.rm = T)`) compared to the principal two (median = `r median(modularityCounts_two, na.rm = T)`, IQR = `r IQR(modularityCounts_two, na.rm = T)`), suggesting that the overall network is composed of two main communities. In addition, the estimation of the optimal number of cuts based on the rank of the first non-zero eigenvalue (see "Partition evaluation" for details) showed that `r prcntDissection` percent of the individual networks are best partitioned into two communities.

@@

``` {r Noise Effects: tSNR Comparison, echo=FALSE, message = FALSE, cache = TRUE}
# Compare SNR between PCC and mPFC per subject after thresholding (this filtering is relevant here because we aren't using ARI anymore)
tempComm_PCC <- lapply(tempComm_PCC, filter, FiedlerVec != 0)
tempComm_PFC <- lapply(tempComm_PFC, filter, FiedlerVec != 0) 
SNR_PCCvsPFC_perm <- sapply(seq_along(tempComm_PCC), function(x) {permute(tempComm_PCC[[x]]$varCoeff, tempComm_PFC[[x]]$varCoeff)$Pval})
SNR_PCCvsPFC_perm <- p.adjust(SNR_PCCvsPFC_perm, method = "BY")
SNR_PCCvsPFC_ES <- sapply(seq_along(tempComm_PCC), function(x) {cohenD(tempComm_PCC[[x]]$varCoeff, tempComm_PFC[[x]]$varCoeff)})
prcntSig <- mean(SNR_PCCvsPFC_perm < 0.05) * 100

@@

### new stuff
# append the Rsquared values onto the current summaries
# unfortunately this only counts on the DMN community
appendVar <- lapply(seq(nSubj), function(subj) {left_join(Summaries[[subj]], Variance[[subj]], by = "Vertex")})

# correlate tSNR with Rsquared before and after thresholds
snrR2_cors <- list()

snrR2_cors$pre_thresh <- sapply(appendVar, function(data) {
  data %>% 
    filter(!is.na(Rsquared)) %>%
    select(varCoeff, Rsquared) %>%
    cor(.$varCoeff)
})[2, ]

snrR2_cors$post_thresh <- sapply(appendVar, function(data) {
  data %>% 
    filter(!is.na(Rsquared), FiedlerBinary != 0.5) %>%
    select(varCoeff, Rsquared) %>%
    cor(.$varCoeff)
})[2, ]

# next get the mean tSNR of PCC and mPFC before/after
snrR2_means <- list()
snrR2_means <- sapply(appendVar, function(data) {
  data %>% 
    filter(!is.na(Rsquared)) %>%
    mutate(region = ifelse(z > 0, "mPFC", "PCC")) %>%
    group_by(region) %>% 
    summarize(mean_tsnr = mean(varCoeff)) %>%
    ungroup()
})

### end new stuff
